---
title       : Pitch For Data Product
subtitle    : Estimating Prevalence of Esophageal Cancer
author      : Herpich
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Motivation

- Esophageal cancer is the eighth most common cancer globally, with nearly 450,000 net cases per year and 400,000 deaths per year

- The most common causes of esophageal cancer, both of ESCC and EAC, include smoking, alcohol, obesity, and poor diet

- Age seems to play a significant impact, as only 15% of cases are found in people younger than 55

- R contains a data frame in the datasets package based on a French case-control study for adults with esophageal cancer in Ille-et-Vilaine

--- .class #id 

## Data

- The esophageal cancer dataset has over 1,100 individual datapoints, both Cases and Controls, which are subcategorized by age group and level of alcohol and tobacco consumption


```{r, echo=FALSE} 
library(datasets)
data(esoph)
unique(esoph$agegp)
unique(esoph$alcgp)
unique(esoph$tobgp)
```

```{r}
sum(esoph$ncontrols) + sum(esoph$ncases)
```

--- .class#id

## Model

- To create a forecasting tool for esophageal cancer prevalance among sub-populations, the data was split into training and test sets, and the training data provided the input for a generalized linear model using 3-fold cross validation

- The predictor variables for this test are the interaction between age and alcohol and age and tobacco as factor variables

- The interaction between tobacco and alcohol proved not statistically significant beyond the original model

- The estimated out-of-sample average MSE was 2% (200 basis points)

```{r,echo=FALSE}
library(caret)
library(datasets)
data(esoph)
esoph$agegp <- unclass(esoph$agegp)
esoph$alcgp <- unclass(esoph$alcgp)
esoph$tobgp <- unclass(esoph$tobgp)
esoph <- cbind(esoph,esoph$ncases/(esoph$ncases+esoph$ncontrols))
names(esoph)[6] <- "Pospercent"
set.seed(1000)
inTrain <- createDataPartition(y=esoph$Pospercent,p=0.9,list=FALSE)
Training <- esoph[inTrain,]
fitControl <- trainControl(method="repeatedcv",number=3,repeats=3)
Model <- train(Pospercent~ (agegp:alcgp) + (agegp:tobgp),data=Training,method="glm",trControl=fitControl)
summary(Model)$coef
```

--- .class#id

## Prediction Tool

- Using the linear model, age, alcohol consumption, and tobacco use can be inputted as factor variables, and the outcome will estimate approximate prevalence of Cases within the population sub-group

- Given the data, age, alcohol and tobacco use are all positive indicators for prevalence of esophageal cancer

- The tool shows the actual data in barchart form so the user can see the relative variation between actual and predicted outcomes

- It should be noted that the data is fairly incomplete, with varying numbers of people in each sub-group and often very few observations (especially in higher age groups); Additionally, it is not clear how biased the population sample is relative to the population as a whole or if there are any confounding variables altering the result, especially since typical case-control studies are self reported by subject

--- .class#id
